---
title: "Simulation experiment"
date: "22 October 2019"
output: pdf_document
---


Define baseline intensity `Z` for the endemic component and an outbreak area `win`.
```{r}
library(spatstat)
temporal.trend.function<-function(t,A=7, B=0.9){
  return (A/2*(sin(t) + 1) + B)
}

spatial.trend.function<-function(x,y, C=80, B=0.5, x0=0.5, y0=0.5){
  return(B + 20*exp(-C * ( (x-x0)^2 + (y - y0)^2)))
}
n.weeks = 20

delta = 0.01
delta2 = delta ^ 2
xx = seq(0,1,delta)
yy= seq(0,1,delta)

Z = lapply(1:n.weeks, function(t)
  outer(xx, yy, function(x,y){temporal.trend.function(t) * spatial.trend.function(x,y)}))

win = owin(poly = data.frame(x=rev(c(0.1,0.2,0.3,0.4,0.44,0.31,0.22)),
                             y=rev(c(0.8,0.91,0.8,0.6,0.55,0.45,0.55))) )

my_blues = c("#F7FBFF", "#DEEBF7", "#C6DBEF", "#9ECAE1", "#6BAED6","#4292C6","#2171B5")
#blues9
ColorRamp=colorRampPalette(c(my_blues))(1000)
```
Generate point events for the baseline and the outbreak with `rpoispp`.
```{r}
set.seed(1)
baseline=lapply(1:n.weeks, function(t){
  rpoispp(function(x,y){ temporal.trend.function(t) * spatial.trend.function(x,y) })})
set.seed(1)
epidemics=lapply(c(0,0,0,0,1,1,1,1,0,0, 0,0,0,0,1,1,1,1,0,0), function(t){
  rpoispp(t * 50, win=win)})

Max = max(unlist(lapply(Z, max)))
```


Plot baseline intensity and point events for each time `t`.
```{r}
# include=FALSE}
# png('../Manuscript/data_points__.png', width = 3.25 *2, height = 3.25 *2, units = 'in', res=600, pointsize = 18)

par(mfrow=c(3,3), mar=c(0.9, 0.9, 0.9, 0.5))
for (i in 1:9){
  ex = max(Z[[i]])/ Max
  ColorRamp_ex = ColorRamp[1:(ex * length(ColorRamp))]

  plot(baseline[[i]], cols='#473B54', pch=16, main=paste0('t=', as.character(i)), show.window = F)
  image(xx, yy, Z[[i]], add=T, col=ColorRamp_ex)
  plot(baseline[[i]], cols='#473B54', pch=16, add=T)
  if ((i == 7) | (i == 8) | (i == 9))
    mtext(c('x' ), c(1), outer = F)
  if ((i == 1) | (i == 4) | (i == 7))
    mtext(c('y' ), c(2), outer = F)
  if (i == 3)
    mtext('A', at = 0.9, cex = 1.8, padj = 1.5)

    
  if (i %in% 5:8){
    plot(win, add=T, border='red')
  }
  plot.ppp(epidemics[[i]], cols='red', pch=16, add=T)
}
# dev.off()
#palette = colorRampPalette(c('blue', 'red'))(max(cases2[,2]))


```

Aggregate all observations in a `list` data structure.
```{r include=FALSE}
observations = list()
for (i in 1:n.weeks){
  b = baseline[[i]]
  s = epidemics[[i]]
  observations[[i]] = data.frame(x=c(b$x, s$x), 
                                 y=c(b$y, s$y),
                                # color=c(rep('#473B54', length(b$x)), rep('red', length(s$x))))
                                 warning=c(rep(F, length(b$x)), rep(T, length(s$x))),
  t=rep(i, length(c(b$x, s$x))))
}
observations = do.call(rbind,  observations)
```

```{r}
writeLines(c("The number of observations is:", nrow(observations)))
```
Define function to draw covering cylinders and compute the exceedance statistics  for each cylinder.
```{r}
# x = seq(0,1,delta)
# y = seq(0,1,delta)
# z = lapply(1:10,
#            function(i){outer(x, y, function(x,y){temporal.trend(i) * spatial.trend.function(x,y)})})


compute_cylinders<-function(cylinder, observations, tabulated.baseline){
  t.low = as.numeric(cylinder['t.low'])
  t.upp = as.numeric(cylinder['t.upp'])
  x0 = as.numeric(cylinder['x'])
  y0 = as.numeric(cylinder['y'])
  rho = as.numeric(cylinder['rho'])
  d = sqrt((tabulated.baseline$x-x0)^2 + (tabulated.baseline$y-y0)^2)
  in_circle = (d < rho) & (!is.na(d))
  in_height = (tabulated.baseline$t >= t.low) & (tabulated.baseline$t <= t.upp)
  mu = sum(tabulated.baseline[in_circle & in_height,]$z) * delta2 #multiply by delta2 as this is the bin size
#   in_of_square = sum(in_circle & in_height) * delta2
# #  out_of_square = pi * rho* rho - in_of_square
#   mu = mu * pi * rho *rho / in_of_square
  
  d = sqrt((observations$x-x0)^2 + (observations$y-y0)^2)
  in_circle = (d < rho) & (!is.na(d))
  in_height = (observations$t >= t.low) & (observations$t <= t.upp)
  n_cases_in_cylinder = sum(in_circle & in_height)
  
 # ci = qpois(c(0.25,0.95) , lambda=mu)
  p.val = ppois(n_cases_in_cylinder, lambda=mu, lower.tail=FALSE)
  return (c(n_cases_in_cylinder, mu, p.val))
}
```

Check timeliness
```{r}
tabulated.baseline = expand.grid(xx,yy,1:n.weeks)
names(tabulated.baseline) = c('x', 'y', 't')
tabulated.baseline$z = apply(tabulated.baseline, 1,
                             function(x){
  temporal.trend.function(x['t']) * spatial.trend.function(x['x'],x['y'])})
# mean.baseline = mean( do.call(rbind, z) )

mean.baseline = mean(tabulated.baseline$z)

total.cases =  nrow(observations)
total.expected.cases = sum(tabulated.baseline$z) * delta2
correction.factor = total.cases / total.expected.cases
tabulated.baseline$z = tabulated.baseline$z * correction.factor
```


```{r}
cylinders.list = list()
for  (t.index in 5:n.weeks){
  n.cylinders = 4000 / n.weeks * t.index
  t.max = min(5, t.index)
  radia = sapply(1:t.max, function(h){sqrt(1 / mean.baseline / pi / h )} )
  radia_and_heights = cbind(1:t.max, radia)

  radia_and_heights = radia_and_heights[sample(1:nrow(radia_and_heights), n.cylinders, replace=T),]
  rho = radia_and_heights[,2]
  random_radia = runif(n.cylinders, 0, rho)
  theta = runif(n.cylinders, 0, 2* pi)
  
  observations.tmp = observations[observations$t <= t.index,]
  tabulated.baseline.tmp = tabulated.baseline[tabulated.baseline$t <= t.index,]
  
  idx = sample(1:nrow(observations.tmp), n.cylinders, replace=T)
  y0 = observations.tmp$y[idx] + sin(theta) * random_radia
  x0 = observations.tmp$x[idx] + cos(theta) * random_radia
  t = observations.tmp$t[idx] + round(runif(n.cylinders, -radia_and_heights[,1]/2, radia_and_heights[,1] / 2))
  
  t.low = t - round(radia_and_heights[,1]/2)
  t.low = ifelse(t.low > 0, t.low, 1)
  t.upp = t + round(radia_and_heights[,1]/2)
  t.max = max(observations.tmp$t)
  t.upp = ifelse(t.upp <= t.max, t.upp, t.max)
  
  cylinders.tmp = data.frame(x=x0, y=y0, rho=rho, t.low=t.low, t.upp=t.upp, original.x=observations.tmp$x[idx],  original.y=observations.tmp$y[idx], original.t=observations.tmp$t[idx])
  
  cylinders.tmp[,c('n_obs', 'mu', 'p.val')] = t(apply(cylinders.tmp, 1, compute_cylinders,
                                                    observations.tmp, tabulated.baseline.tmp))
  
 # cylinders.tmp$warning = apply(cylinders.tmp, 1, function(x){ifelse((x['p.val'] < 0.05) & (x['n_obs'] > 0), TRUE, FALSE)})
  
  cylinders.tmp$warning = (cylinders.tmp$p.val < 0.05) & (cylinders.tmp$n_obs > 0)
  
  cylinders.list[[t.index]] = cylinders.tmp
}
rm(cylinders.tmp)
rm(tabulated.baseline.tmp)
rm(observations.tmp)
```


Compute the warning scores.

```{r}
warning.score<-function(observation, cylinders){
  # check if the location
  x = as.numeric(observation['x'])
  y = as.numeric(observation['y'])
  TT = as.numeric(observation['t'])
  in_circle = as.integer(sqrt((as.numeric(cylinders$x) - x)^2 + (as.numeric(cylinders$y) - y)^2) < as.numeric(cylinders$rho))
  in_cylinder_height = as.integer( (as.numeric(cylinders$t.low) <= TT) & (as.numeric(cylinders$t.upp) >= TT))
  
  # number of cylinders that include geo-coordinate of `case`
  in_cylinder = sum(in_circle * in_cylinder_height, na.rm=T)

  # number of cylinder with `warning` flag that include location `i`
  warning = sum(cylinders$warning * in_circle * in_cylinder_height,na.rm=T)
  if (in_cylinder>0){
    re = warning / in_cylinder
  }else{
    re = 0
  } 
  return(re)
}
# init = Sys.time()
# observations$warning.score2 = apply(observations, 1, warning.score, cylinders)
# print(Sys.time() - init)
```

Timeliness warning scores
```{r}
observations.list = list()
for (t.index in 5:n.weeks){
  observations.list[[t.index]] = observations[observations$t <= t.index,]
}
```

```{r}
for (t.index in 5:n.weeks){
  observations.list[[t.index]]$warning.score = apply(observations.list[[t.index]], 1, warning.score, cylinders.list[[t.index]])
}
```



<!-- ```{r} -->
<!-- #png('../Fig/timeliless_25.png', res=600, width=5, height=4, units="in") -->
<!-- plot(c(1,n.weeks), c(0.2,1), type = 'n', ylab = 'AUC', xlab = expression(tau)) -->
<!-- # -->
<!-- cols = c("#1f77b455", "#ff7f0e55", "#2ca02c55", "#d6272855", "#9467bd55", "#8c564b55", "#e377c255", "#7f7f7f55", "#bcbd2255", "#17becf55") -->

<!-- x = c(5, 6, 7, 8, 15, 16, 17, 18) -->
<!-- for (I in 1:length(x)){ -->
<!--   aucs = 0 -->
<!--   t.index = x[I] -->
<!--   xx = 0 -->
<!--   for (tau in t.index:n.weeks){ -->
<!--     idx = observations.list[[tau]]$t == t.index -->
<!--     if (length(unique(observations.list[[tau]][idx,]$warning))>1){ -->
<!--       cc = pROC::roc(as.integer(warning) ~ warning.score,  observations.list[[tau]][idx,]) -->
<!--       cc = cc$auc -->
<!--       aucs =c(aucs, cc) -->
<!--       xx = c(xx, tau) -->
<!--     } -->
<!--   } -->
<!--   print(c(x[I], length(xx))) -->
<!--   lines(xx[-1], aucs[-1], col=cols[I], lwd=3) -->
<!-- } -->
<!-- legend("bottomleft",legend = sapply(x, function(x){paste0("t=", as.character(x))}), col=cols[1:length(x)], lwd = 3) -->
<!-- # dev.off() -->
<!-- ``` -->

<!-- ``` -->
<!-- init_size = observations.list[[20]]$t %>% table %>% as.data.frame() -->
<!-- for (t in 5:19){ -->
<!--   idx = init_size$. == t -->
<!--   if (sum(idx) > 0){ -->
<!--     ws= matrix(ncol = init_size$Freq[idx]) -->
<!--     idx = observations.list[[20]]$t == t -->
<!--     color = ifelse(observations.list[[20]][idx,]$warning, 'red', 'blue') -->

<!--     for (tau in t:20){ -->
<!-- #      idx = observations.list[[tau]]$t == t -->
<!--       w = observations.list[[tau]][observations.list[[tau]]$t == t,]$warning.score -->
<!--       ws = rbind(ws, w) -->
<!--     } -->
<!--     ws = ws[-1,] -->

<!--     if (!is.matrix(ws)){ -->
<!--       ws = matrix(ws, nrow=length(ws)) -->
<!--     } -->

<!--     # png(paste0('../Manuscript/timeliness_', as.character(t) , '.png'), width = 5, height = 4, units = 'in', res=600, pointsize = 10) -->
<!--     plot(c(t,t+nrow(ws)-1),range(ws), type='n', xlab = expression(tau), ylab=expression(italic('w')), main = paste0('t=',as.character(t))) -->
<!--     for(i in 1:ncol(ws)){ -->
<!--       lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty =  2) -->
<!--       points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = 19) -->
<!--       legend('bottomright', c('outbreak','endemic'), pch=19, col=c('red', 'blue')) -->
<!--     } -->
<!--     # dev.off() -->
<!--   } -->
<!-- } -->
<!-- ``` -->


```{r}
tab.red = '#d62728'
tab.blue = '#1f77b4'
init_size = as.data.frame(table(observations.list[[20]]$t ))
#png(paste0('../Manuscript/timeliness_','.png'), width = 5 * 2, height = 4, units = 'in', res=600, pointsize = 12)
par(mfrow=c(1, 2)) #, mar=c(0.9,0.9,0.9,0.7))
plot(
  c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(italic(w)), axes=FALSE)
for (t in 5:6){
  # idx = init_size$. == t
  idx = init_size$Var1 == t  
  if (sum(idx) > 0){
    ws= matrix(ncol = init_size$Freq[idx])
    idx = observations.list[[20]]$t == t
    color = ifelse(observations.list[[20]][idx,]$warning, tab.red, tab.blue)
    
    for (tau in t:20){
      idx = observations.list[[tau]]$t == t
      w = observations.list[[tau]][idx,]$warning.score
      ws = rbind(ws, w)
    }
    ws = ws[-1,]
    
    if (t==5){
      pch=15
      lty=2
    }else if(t==6){
      pch=17
      lty=3
    }
    
    for(i in 1:ncol(ws)){
      lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty =  lty)
      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = pch, cex=0.5)
    }
    axis(side = 1, at = 5:20)
    axis(side = 2)
  }
}
legend('bottomright', c('outbreak, t=5','endemic, t=5', 'outbreak, t=6','endemic, t=6'),
       lty=c(2,2,3,3), lwd=c(1,1,1,1),
       col=c(tab.red, tab.blue, tab.red, tab.blue),
       pch=c(15,17,17,17), pt.cex=0.5)
# legend('bottom', c('t=5','t=6'), pch=c(5,6) )
plot(
  c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(italic(w)), axes=FALSE)
for (t in 7){
  # idx = init_size$. == t
  idx = init_size$Var1 == t  
  if (sum(idx) > 0){
    ws= matrix(ncol = init_size$Freq[idx])
    idx = observations.list[[20]]$t == t
    color = ifelse(observations.list[[20]][idx,]$warning,  tab.red, tab.blue)
    
    for (tau in t:20){
      idx = observations.list[[tau]]$t == t
      w = observations.list[[tau]][idx,]$warning.score
      ws = rbind(ws, w)
    }
    ws = ws[-1,]
    for(i in ncol(ws):1){
      lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty =  2)
      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = 1, cex = 0.5)
    }
    axis(side = 1, at = 5:20)
    axis(side = 2)
  }
}
legend('bottomright', c('outbreak, t=7','endemic, t=7'), lwd=c(1,1,NA),
       lty=c(1,1), pch=c(1,1), col=c( tab.red, tab.blue), pt.cex=0.5)
# legend('bottom', c('t=7'), pch=c(1) )
# dev.off()
```


```{r}
tab.red = '#d62728'
tab.blue = '#1f77b4'
init_size = as.data.frame(table(observations.list[[20]]$t ))
# png(paste0('../Manuscript/timeliness__','.png'), width = 5 * 2.8, height = 4, units = 'in', res=600, pointsize = 20)
par(mfrow=c(1, 3), mar= c(4, 4, 0.5, 0.5) )
plot(
  c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(italic(w)), axes=FALSE, cex.lab=1.4)
t = 5
  # idx = init_size$. == t
  idx = init_size$Var1 == t  
  if (sum(idx) > 0){
    ws= matrix(ncol = init_size$Freq[idx])
    idx = observations.list[[20]]$t == t
    color = ifelse(observations.list[[20]][idx,]$warning, tab.red, tab.blue)
    pch = ifelse(observations.list[[20]][idx,]$warning, 15, 17)
    lty = ifelse(observations.list[[20]][idx,]$warning, 1, 2)
    
    for (tau in t:20){
      idx = observations.list[[tau]]$t == t
      w = observations.list[[tau]][idx,]$warning.score
      ws = rbind(ws, w)
    }
    ws = ws[-1,]

    for(i in 1:ncol(ws)){
      lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty =  lty[i])
      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = pch[i], cex=0.5)
    }
    axis(side = 1, at = 5:20)
    axis(side = 2)
  }
abline(h=0.95, col='#7f7f7f')
legend('bottomright', c('outbreak, t=5','endemic, t=5'), lwd=c(1,1),
       lty=c(1,2), col=c(tab.red, tab.blue), pch=c(15,17), pt.cex=0.5)
plot(
  c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(italic(w)), axes=FALSE, cex.lab=1.4)
t = 6
  # idx = init_size$. == t
  idx = init_size$Var1 == t  
  if (sum(idx) > 0){
    ws= matrix(ncol = init_size$Freq[idx])
    idx = observations.list[[20]]$t == t
    color = ifelse(observations.list[[20]][idx,]$warning, tab.red, tab.blue)
    pch = ifelse(observations.list[[20]][idx,]$warning, 15, 17)
    lty = ifelse(observations.list[[20]][idx,]$warning, 1, 2)

    for (tau in t:20){
      idx = observations.list[[tau]]$t == t
      w = observations.list[[tau]][idx,]$warning.score
      ws = rbind(ws, w)
    }
    ws = ws[-1,]
    
    for(i in 1:ncol(ws)){
      lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty = lty[i])
      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = pch[i], cex=0.5)
    }
    axis(side = 1, at = 5:20)
    axis(side = 2)
  }
abline(h=0.95, col='#7f7f7f')
legend('bottomright', c('outbreak, t=6','endemic, t=6'), lwd=c(1,1),
       lty=c(1, 2), col=c(tab.red, tab.blue),  pch=c(15,17), pt.cex=0.5)
plot(
  c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(italic(w)), axes=FALSE,  cex.lab=1.4)
  t=7
  # idx = init_size$. == t
  idx = init_size$Var1 == t  
  if (sum(idx) > 0){
    ws= matrix(ncol = init_size$Freq[idx])
    idx = observations.list[[20]]$t == t
    color = ifelse(observations.list[[20]][idx,]$warning,  tab.red, tab.blue)
    pch = ifelse(observations.list[[20]][idx,]$warning, 15, 17)
    lty = ifelse(observations.list[[20]][idx,]$warning, 1, 2)

    for (tau in t:20){
      idx = observations.list[[tau]]$t == t
      w = observations.list[[tau]][idx,]$warning.score
      ws = rbind(ws, w)
    }
    ws = ws[-1,]
    for(i in ncol(ws):1){
      lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty = lty[i])
      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = pch[i], cex = 0.5)
    }
    axis(side = 1, at = 5:20)
    axis(side = 2)
  }
abline(h=0.95, col='#7f7f7f')
legend('bottomright', c('outbreak, t=7','endemic, t=7'), lwd=c(1, 1),
       lty=c(1,2), pch=c(15,17), col=c(tab.red, tab.blue), pt.cex=0.5)

# dev.off()
```
<!-- ```{r} -->
<!-- init_size = observations.list[[20]]$t %>% table %>% as.data.frame() -->
<!-- png(paste0('../Manuscript/timeliness_2','.png'), width = 5, height = 4, units = 'in', res=600, pointsize = 10) -->
<!-- par(mfrow=c(1, 1)) #, mar=c(0.9,0.9,0.9,0.7)) -->
<!-- plot( -->
<!--   c(5,20), c(0,1), type='n', xlab = expression(tau), ylab=expression(bold('w'))) -->
<!-- for (t in 5:7){ -->
<!--   idx = init_size$. == t -->
<!--   if (sum(idx) > 0){ -->
<!--     ws= matrix(ncol = init_size$Freq[idx]) -->
<!--     idx = observations.list[[20]]$t == t -->
<!--     color = ifelse(observations.list[[20]][idx,]$warning, 'red', 'blue') -->

<!--     for (tau in t:20){ -->
<!--       idx = observations.list[[tau]]$t == t -->
<!--       w = observations.list[[tau]][idx,]$warning.score -->
<!--       ws = rbind(ws, w) -->
<!--     } -->
<!--     ws = ws[-1,] -->
<!--     for(i in 1:ncol(ws)){ -->
<!--       lines(t:(t+nrow(ws)-1), ws[,i], col=color[i], lty =  t -3) -->
<!-- #      points(t:(t+nrow(ws)-1), ws[,i], col=color[i], pch = t) -->
<!--     } -->
<!--   } -->
<!-- } -->
<!-- legend('bottomright', c('outbreak','endemic', 't=5','t=6', 't=7'), lty=c(1,1,2,3,4), col=c('red', 'blue', 'black', 'black', 'black')) -->
<!-- dev.off() -->
<!-- ``` -->




<!-- ``` -->
<!-- # png('../Manuscript/prediction_points_7.png', width = 3.25 *2, height = 3.25 *2, units = 'in', res=600, pointsize = 18) -->
<!-- par(mfrow=c(3,3), mar=c(0.9,0.9,0.9,0.7)) -->
<!-- observations[,"warning.color"] = round(observations[,"warning.score2"]  * 10) + 1  -->
<!-- #library(viridis) -->
<!-- #palette = viridis(11, begin = 0, end = 1) -->
<!-- palette = colorRampPalette(c("blue", "red"))(11) ## (n) -->
<!-- win2 = owin(poly = data.frame(x=c(0,1,1,0),y=c(0,0,1,1)) ) -->
<!-- for (i in 1:9){ -->
<!--   idx = (observations$t == i) -->
<!--   colors = palette[observations[idx,"warning.color"]] -->
<!--   PPP = ppp(observations[idx,"x"], observations[idx,"y"], window = win2) -->
<!--   plot(PPP, cex=0.1,  pch=16, main=paste0('t=', as.character(i)), show.window = F) -->

<!--   ex = max(Z[[i]]) / Max -->
<!--   ColorRamp_ex = ColorRamp[1:(ex * length(ColorRamp))] -->
<!--   image(xx, yy, Z[[i]], add=T, col=ColorRamp_ex) -->
<!-- #  plot(PPP, cex=0.1,  pch=16). -->

<!--   if ((i == 7) | (i == 8) | (i == 9)) -->
<!--     mtext(c('x' ), c(1), outer = F) -->
<!--   if ((i == 1) | (i == 4) | (i == 7)) -->
<!--     mtext(c('y' ), c(2), outer = F) -->
<!--   if (i == 3) -->
<!--     mtext('B', at = 0.9, cex = 1.8, padj = 1.5) -->

<!--   points(observations[idx,"x"], observations[idx,"y"], col=colors, pch=16, cex=1) -->
<!--   if (i %in% 5:8){ -->
<!--     plot(win, add=T, border='red') -->
<!--   } -->
<!-- } -->
<!-- # dev.off() -->
<!-- ``` -->


<!-- ``` -->
<!-- cc1 = pROC::roc(as.integer(warning) ~ warning.score, observations) -->
<!-- cc2 = pROC::roc(as.integer(warning) ~ warning.score2, observations) -->

<!-- df1 = data.frame(cc1$specificities, cc1$sensitivities, cc1$thresholds) -->
<!-- df2 = data.frame(cc2$specificities, cc2$sensitivities, cc2$thresholds) -->

<!-- idxa = which(df1$cc1.thresholds > 0.95) -->
<!-- idxb = which(df1$cc1.thresholds < 0.95) -->
<!-- idx = c(idxa[1], idxb[length(idxb)]) -->
<!-- # print(df1) -->
<!-- spec_sens_cc1 = colMeans(df1[idx,]) -->


<!-- idxa = which(df2$cc2.thresholds > 0.95) -->
<!-- idxb = which(df2$cc2.thresholds < 0.95) -->
<!-- idx = c(idxa[1], idxb[length(idxb)]) -->
<!-- # print(df2) -->
<!-- spec_sens_cc2 = colMeans(df2[idx,]) -->


<!-- ``` -->

<!-- Sensitivity: the ability of a test to correctly identify epidemic events. -->
<!-- Specificity: the ability of a test to correctly identify non-epidemic events. -->


<!-- ```{r} -->
<!-- options(warn = - 1) -->
<!-- #conf_matrix<-table(ifelse(observations$warning.score2 > 0.95, T, F), observations$warning) -->
<!-- # sens = caret::sensitivity(conf_matrix) -->
<!-- # spec = caret::specificity(conf_matrix) -->
<!-- # cat("sensitivity=", sens, "\n") -->
<!-- # cat("specificity=", spec, "\n") -->
<!-- # png('../Manuscript/ROC_.png', width = 3.25, height = 3.25, units = 'in', res=400, pointsize = 13) -->
<!-- par(mfrow=c(1,1), mar=c(0.5, 0.5, 0.5, 0.5)) -->
<!-- col = '#00000033' -->
<!-- cc = pROC::roc(as.integer(warning) ~ warning.score2, observations) -->
<!-- plot(cc, col='black', identity.col='black',  grid.col='red', grid.v=spec_sens_cc1[1], grid.h = spec_sens_cc1[2]) -->
<!-- auc = as.numeric(cc['auc'][[1]]) -->
<!-- L = nrow(observations) -->
<!-- for (i in 1:50){ -->
<!--   cc = pROC::roc(as.integer(warning) ~ warning.score2, observations[sample(1:L, L, replace = T),]) -->
<!--   if(i < 11){ -->
<!--     plot(cc, add=T, col=col,  identity.col='black')     -->
<!--   } -->
<!--   auc = c(auc,as.numeric(cc['auc'][[1]])) -->
<!-- } -->
<!-- print(quantile(auc, c(0.025,0.5,0.975) )) -->
<!-- #     2.5%       50%     97.5%  -->
<!-- # 0.9567308 0.9792924  -->
<!-- # 0.9279978 0.9522443 0.9749876 -->
<!-- # dev.off() -->
<!-- ``` -->

